---
title: "Data Visualisation"
author: "Hans"
date: "9/10/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) +
  install.packages("pacman") 
library(pacman)
```


```{r}
p_load(tidyverse)
p_load(gapminder)
theme_set(theme_minimal())
```

```{r}
glimpse(gapminder)
```

First set up the frame in which to plot

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p
```


Then add the code to create a dot plot

```{r}
p + geom_point()
```

Or, rather than a dot plot, get a smooth line

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth(se = F)
```

If you like, add a confidence interval around the line 

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth(se = F)
```

Or have both the points and the line!

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth() + 
  geom_point()
```


Rather than a smooth line, you can also draw a regression line

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth(method = "lm") + 
  geom_point()
```

Make the x-axis logarithmic

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth(method = "lm") + 
  geom_point() + 
  scale_x_log10()
```

Format the numbers on the x-axis as dollars

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth(method = "lm") + 
  geom_point() + 
  scale_x_log10(labels=scales::dollar)
  
```


Now use color to visually separate the continents 

```{r}
p <- ggplot(data=gapminder, mapping = aes(x = gdpPercap, y = lifeExp, 
                                          color = continent, fill = continent))
p + geom_smooth() + 
  geom_point() + 
  scale_x_log10(labels=scales::dollar)
  
```

You can apply te color only to the points and not to the smooth line

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth() + 
  geom_point(mapping = aes(color = continent, fill = continent)) + 
  scale_x_log10(labels=scales::dollar) 
```

Add some decent looking labels

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth() + 
  geom_point(mapping = aes(color = continent, fill = continent)) + 
  scale_x_log10(labels=scales::dollar) +
  labs(title = "Plot Title\n", x = "GDP per capita", 
       y = "Life Expectancy", colour = "Continents\n")
```

What if you want to make the points less dominant? Use Alpha

```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth() + 
  geom_point(mapping = aes(color = continent, fill = continent, alpha = 0.3)) + 
  scale_x_log10(labels=scales::dollar) +
  labs(title = "Plot Title\n", x = "GDP per capita", 
       y = "Life Expectancy", colour = "Continents\n")
```

It can be simplified; you do not need to ie the variable names 'data' and 'mapping' 

```{r}
p <- ggplot(gapminder,
            aes(x = gdpPercap, y = lifeExp))
p + geom_smooth() + 
  geom_point(aes(color = continent, fill = continent, alpha = 0.3)) + 
  scale_x_log10(labels=scales::dollar) +
  labs(title = "Plot Title\n", x = "GDP per capita", 
       y = "Life Expectancy", colour = "Continents\n")
```


