---
title: "Plotting methods"
author: "Hans"
date: "9/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse) 
```


```{r}
library(deSolve)

expgrowth = function(timepoint, state, parameters) { 
  with (as.list(c(state, parameters)), {
    dY = a * Y
    list(c(dY)) 
  })
}

pars = c(a = -0.1)
inistate1 = c(Y = 1)
inistate2 = c(Y = 5)
inistate3 = c(Y = 10)
time_series = seq(0, 10, by = 0.1)

# Here you find the solutions for the ODE abd generate the datasets
# The dataset contains an Y and a Time column
out1 = ode(y = inistate1, times = time_series, func = expgrowth, parms = pars)
out2 = ode(y = inistate2, times = time_series, func = expgrowth, parms = pars)
out3 = ode(y = inistate3, times = time_series, func = expgrowth, parms = pars)
```

# Method 1 - Plot with empty plot, then add lines

Plotting objects in an empty plot, using the lines command

First create an empty plot, by using type = 'n'
You also specify a dummy parameter x (=1)

```{r}
plot(1, xlab = "time", ylab = "Y ", type = "n", 
     xlim = c(0, 10), 
     ylim = c(min(out3), max(out3)))
 
# Because out1 already contains all the data, you do not need to specify y data
lines(out1, ylab = "Y = 1", col = "blue", lty=2)
lines(out2, ylab = "Y = 5", col = "red")
lines(out3, ylab = "Y = 10", col = "black")
```


# Method 2 - Add datasets directly to plot

You can simply provide the tree datsets to plot

```{r}
plot(out1, out2, out3)
```


# Method 3 - Do first line with plot, then add lines

You can plot the first data set with plot and then add lines for the other datasets.

That gives youbthe ability to control the individual lines better

```{r}
plot(out1, main = "Plot Title", ylab = "Y", col = "red", ylim = c(0, 10), lty=4) 
lines(out2, col = "blue", lty=2)
lines(out3, col = "green", lty=3)
```


# Method 4 Different approach: plot with the curve command

ggplot expects a dataframe, so you have tgo convert

```{r}
rdata = as.data.frame(out1)
ggplot(data = rdata) +
  geom_line(mapping = aes(x = time, y = Y), color='blue')
```

```{r}

decay <- function(t, k, P, A){
  A * exp(-k*t) * sin (2*pi*t/P)
}

# Here you provide a function to plot
curve(decay(x, 0.1, 4, 1), 0, 50, xlab="t", ylab="decay", col="red", n = 1000)
```
